\SW: Crowd-sourcing the Discovery of Gravitational Lenses

### 2014-05-07 doc/sw-system.tex
Finished SWAP section and appendix

\title[\sw I]
{\SW: I. Crowd-sourcing the Discovery of Gravitational Lenses}

@@

\sw is a web-based service that enables the discovery of strong gravitational
lenses in wide-field imaging surveys by large numbers of people. Carefully
produced color composite images are displayed to volunteers via a flexible
interface, which records their estimates of the positions of candidate lensed
features. Simulated lenses, and expert-classified non-lenses, are inserted
into the stream at random intervals; this training set is used to give the
volunteers feedback on their performance, and to estimate a
dynamically-updated probability for any given image to contain a lens. Low
probability systems are retired from the site daily, concentrating the sample
towards a set of candidates; these are then re-classified by the volunteers in
a second  refinement stage. Analyzing the classification of the training set,
we predict that the first stage should yield a sample with C\% completeness
and P\% purity, while leading to the rejection of R\% of the initial target
sample. Having divided the 150 square degree CFHTLS imaging survey into 430000
overlapping 70 by 70 arcminute tiles and displayed them on the site, we were
joined by 33000 volunteers who contributed X million image classifications
over the course of N months. The sample was reduced to 3500 stage 1
candidates; these were then refined to yield a sample of candidates rankable
by their stage 2 probability. We expect this sample to be X\% complete and Y\%
pure at a threshold of 95\% classification probability. We estimate the mean
information contributed per person to be X bits, over a session lasting, on
average, N classifications per volunteer -- although the distributions of
these quantities are highly skewed. We comment on the scalability of the \sw
system, and its potential to operate beyond its design as a supervised
classification system.

@@

\section{Classification Analysis}
\label{sec:IDanalysis}

@@

We perform a ``Pseudo-online'' analysis of the classifications, 
updating a probabilistic model of every (anonymous) volunteer's
data, and also updating the lens probability of each subject 
(in both the training and test sets), on a
daily(??) basis. This allows us to track the speed with which the crowd
learns about lenses, and also gives us a dynamic estimate of the
posterior probability for  any given  subject being a lens, given all
classifications of it. Assigning thresholds in this lens probability
allows us to make good decisions about whether or not to accept a
subject into the collection of candidates visible in \Talk, and also
whether or not to  carry on classifying a subject at all.

In this paper we focus on the training data,  investigating  how the 
crowd's ability to identify gravitational lenses during the course of
the project, and the completeness and purity of the lens candidate
sample generated.

Describe SWAP here. (Move from appendix.)

@@

\item Completeness and purity were estimated as C\% and P\%, from sim and dud
recovery/miss rates Which sims were missed? False negatives

@@

\section{Probabilistic Classification Data Model}
\label{appendix:probmodel}

Our aim is to enable the construction of a sample of good lens candidates.
Since we aspire to making logical  decisions, we define a  ``good candidate''
as one which has a high posterior probability of being a lens, given the data:
$\pr(\LENS|\data)$. Our problem is to approximate this probability. The data
in our case are the pixel values of a colour image. However, we can greatly
compress these complex, noisy sets of data by asking each volunteer what they
think about them. A complete  classification in \sw consists of a set of
Marker positions, or none at all. The null set encodes the statement from
the volunteer that the image in question is $\saidNOT$ a lens, while the
placement of any  Markers indicates that the volunteer considers this image to
contain a $\saidLENS$.  We simplify the problem by only using the Marker
positions to assess whether t
he volunteer  correctly assigned the
classification $\saidLENS$ or $\saidNOT$ after viewing (blindly) a member of
the training set of subjects.

How should we model these compressed data? The circumstances of each
classification are quite complex: the volunteers learn more about the problem
as they go, but also inevitably make occasional mistakes (perhaps because a
lens is difficult to see, or they became distracted by the television). To
cope with this uncertainty, we assign a software agent to partner each
volunteer. The agent's task is to interpret their volunteer's classification
data as best it can, using a model that makes a number of necessary
approximations. These interpretations will then include uncertainty arising as
a result of the volunteer's efforts and also the agent's approximations, but
they will have two important redeeming features. First, the interpretations
will be quantitative (where before they were qualititative),  and thus will be
useful in decision-making. Second, the agent will be able to predict, using
its model, the probability of a test subject being a $\LENS$, given both its
volunteer's classification, and its volunteer's experience. We now
describe how each agent works.

@@

As an example, suppose we have a volunteer who is always right about the true
nature of a training subject. 
Their agent's confusion matrix would be
\be
  \mathcal{M}^{\rm perfect} = 
  \begin{bmatrix}
    0.0 & 1.0 \\
    1.0 & 0.0
  \end{bmatrix}.
\ee
On being given a fresh subject that actually is a $\LENS$, this hypothetical
volunteer would submit $C = \saidLENS$.  Their agent would then calculate the
posterior probability for the subject being a $LENS$ to be
\begin{align}
  \pr(\LENS|\saidLENS,T_k) &= \frac{1.0 \cdot p_0}
           {\left[ 1.0\cdot p_0 + 0.0\cdot(1 - p_0) \right]}
   &= 1.0,
\end{align}
as we might expect for such a {\it perfect} classifier.  Meanwhile, a
hypothetical volunteer who (for some reason) wilfully always submits the wrong
classification would have an agent with the column-swapped confusion matrix
\be
  \mathcal{M}^{\rm obtuse} = 
  \begin{bmatrix}
    1.0 & 0.0 \\
    0.0 & 1.0
  \end{bmatrix},
\ee
and would submit $C = \saidNOT$ for this subject. However, such a volunteer
would nevertheless be submitting useful information, since given the above
confusion matrix, their agent would calculate
\begin{align}
  \pr(\LENS|\saidNOT,T_k) &= \frac{1.0 \cdot p_0}
           {\left[ 1.0\cdot p_0 + 0.0\cdot(1 - p_0) \right]}
   &= 1.0.
\end{align}
{\it Obtuse} classifiers are as helpful as {\it perfect} ones!

Indeed, the information content of each classification can be estimated by an
agent before the next classification is made, just from its confusion matrix.
The Shannon entropy generated by a classifier upon performing a
classification is
\be
  \langle S_k \rangle =  
   -P_{\rm right}\cdot\log_2{P_{\rm right}} -  P_{\rm wrong}\cdot\log_2P_{\rm wrong}\,,
\ee
where $P_{\rm right}$ and $P_{\rm wrong}$ are the averages of the diagonal and
the off-diagonal elements of the confusion matrix, respectively, and $\langle
S_k \rangle$ is measured in ``bits.'' These averages represent the probability
of a classifier to get a classification right or wrong, respectively. We define
the information contributed by a classifier as
\begin{equation}
I_k = 1 - S_k\,.
  \label{eq:app:info}
\end{equation}
\Eref{eq:app:info} gives the required result, that both the hypothetical {\it
perfect} and {\it obtuse} classifiers contribute 1 bit of information each, per
classification. Classifiers whose agent's confusion matrix is such that $P_{\rm
right}=P_{\rm wrong}=0.5$, contribute zero bits of information. Such users
identify a lens correctly with the same probability as they misclassify a dud
image to contain a lens, and thus their classification is of no value.

@@

@@

The confusion matrix obtained from the application of Equation
(\ref{eq:app:fraction}) has some inherent noise which reduces as the number
of training subjects classified by the user increases. For simplicity, the
discussion thus far assumed the case when the confusion matrix is known
perfectly. Let us first discuss how to characterize the noise in the confusion
matrix. ** This needs work **

For ease of notation, we will denote $\pr(C_k|\LENS,T_k)\equiv p_L$ and
$\pr(C_k|\NOT,T_k)\equiv p_N$. In reality, there is a probability distribution
for both $p_L$ and $p_N$. Let $p_0$ be the prior probability of the subject
being a lens. Then the posterior probability, $p_0'$ of the subject being a
lens after the classification $C_k$ is
\be
  \label{eq:app:sec}
p_0' = \frac{p_L p_0}{\left[ p_L p_0 + p_N (1- p_0) \right]},
\ee
The posterior probability distribution $p_0'$ can be obtained by marginalizing
over the probability distributions of $p_L$, $p_N$ and the prior probability
distribution $p_0$ such that,
\be
P(p_0') = \int p_0' P(p_L) P(p_N) P(p_0) dp_L dp_N dp_0\,.
\ee
This marginalization is not analytically tractable. Therefore, we have
implemented the following Monte-Carlo solution for this problem.

>>>>

\title[\sw I]
{\SW: Crowd-sourcing the Discovery of Gravitational Lenses}

@@

\sw is a web-based service that enables the discovery of strong gravitational
lenses in wide-field imaging surveys by large numbers of people. Carefully
produced color composite images are displayed to volunteers via a
classification interface which records their estimates of the positions of
candidate lensed features. Simulated lenses, and expert-classified non-lenses,
are inserted into the image stream at random intervals; this training set is
used to give the volunteers feedback on their performance, and to estimate a
dynamically-updated probability for any given image to contain a lens. Low
probability systems are retired from the site periodically, concentrating the
sample towards a set of candidates; this ``stage 1'' set is then re-classified
by the volunteers in a second  refinement stage. Analyzing the classification
of the training set, we predict that the first stage alone should yield a
sample that is C\% complete, while leading to the rejection of R\% of the
initial target sample. Having divided the 150 square degree CFHTLS imaging
survey into 430000 overlapping 70 by 70 arcminute tiles and displayed them on
the site, we were joined by 33000 volunteers who contributed X million image
classifications over the course of N months. The sample was reduced to 3500
stage 1 candidates; these were then refined to yield a sample of 1400
candidates rankable by their stage 2 probability. We expect this sample to be
X\% complete and Y\% pure at a threshold of 95\% classification probability.
We find that, on average, and given the assumptions we make in our analysis,
we need 9 classifications per image during the first stage, X in the second.
We estimate the mean information contributed per person to be X bits, over a
session lasting, on average, N classifications per volunteer, and present the
highly skewed distributions of these quantities. We comment on the scalability
of the \sw system to the wide field survey era, and its potential to operate
beyond its design as a supervised classification system.

@@

\section{Classification Analysis}
\label{sec:swap}

@@

We perform an online analysis of the classifications,  updating a
probabilistic model of every (anonymous) volunteer's data, and also updating
the lens probability of each subject  (in both the training and test sets), on
a daily basis. This gives us a dynamic estimate of the posterior probability
for  any given  subject being a lens, given all classifications of it to date.
Assigning thresholds in this lens probability then allows us to make good
decisions about whether or not to retire a subject from the system, in order to
focus attention on new images.

The details of how the lens probabilities are calculated are given in
\Aref{appendix:swap}. In summary:
\begin{itemize}

\item Each volunteer is assigned a simple software agent, characterised by a
confusion matrix. The two independent elements of this matrix are the
probabilities, as estimated by the agent, that the volunteer is going to be 1)
correct when they report that an image contains a lens when it really does
contain a lens, $\pr(\saidLENS|\LENS,T)$, and 2) correct when they report that
an image does not contain a lens when it really doesn't contain a lens,
$\pr(\saidNOT|\LENS,T)$.

\item Each agent updates its confusion matrix elements based on the number of
times its volunteer has been right in each way while classifying subjects from
the training set, accounting for noise early on due to small number
statistics: $T$ is the set of all training images seen to date.

\item Each agent uses its confusion matrices to update, via Bayes' theorem,
the probability of an image from the test set containing a lens,
$\pr(\LENS|C,T)$, when that image is classified by its volunteer. ($C$ is the
set of all classifications made of this subject.)

In the next section we present results in terms of all three of these
probabilities, as we investigate the performance of the \SW system.

@@

\item Completeness and purity were estimated as C\% and P\%, from sim and dud
recovery/miss rates. Which sims were missed? False negatives

@@

\section{Probabilistic Classification Analysis}
\label{appendix:swap}

Our aim is to enable the construction of a sample of good lens candidates.
Since we aspire to making logical  decisions, we define a  ``good candidate''
as one which has a high posterior probability of being a lens, given the data:
$\pr(\LENS|\data)$. Our problem is to approximate this probability. The data~$\data$
in our case are the pixel values of a colour image. However, we can greatly
compress these complex, noisy sets of data by asking each volunteer what they
think about them. A complete  classification in \sw consists of a set of
Marker positions, or none at all. The null set encodes the statement from
the volunteer that the image in question is $\saidNOT$ a lens, while the
placement of any  Markers indicates that the volunteer considers this image to
contain a $\saidLENS$.  We simplify the problem by only using the Marker
positions to assess whether the volunteer  correctly assigned the
classification $\saidLENS$ or $\saidNOT$ after viewing (blindly) a member of
the training set of subjects.

How should we model these compressed data? The circumstances of each
classification are quite complex, as are the human classifiers in general: the
volunteers learn more about the problem as they go, but also inevitably make
occasional mistakes (perhaps because a lens is difficult to see, or they
became distracted during the task). To cope with this uncertainty, we assign a
simple software {\it agent} to partner each volunteer. The agent's task is to
interpret their volunteer's classification data as best it can, using a model
that makes a number of necessary approximations. These interpretations will
then include uncertainty arising as a result of the volunteer's efforts and
also the agent's approximations, but they will have two important redeeming
features. First, the interpretations will be quantitative (where before they
were qualititative),  and thus will be useful in decision-making. Second, the
agent will be able to predict, using its model, the probability of a test
subject being a $\LENS$, given both its volunteer's classification, and its
volunteer's experience. In this appendix we describe how these agents work.

\subsection{Agents and their Confusion Matrices}
\label{appendix:swap:probabilities}

@@

\subsection{Examples}
\label{appendix:swap:examples}

Suppose we have a volunteer who is always right about the true
nature of a training subject. 
Their agent's confusion matrix would be
\be
  \mathcal{M}^{\rm perfect} = 
  \begin{bmatrix}
    0.0 & 1.0 \\
    1.0 & 0.0
  \end{bmatrix}.
\ee
On being given a fresh subject that actually is a $\LENS$, this hypothetical
volunteer would submit $C = \saidLENS$.  Their agent would then calculate the
posterior probability for the subject being a $LENS$ to be
\begin{align}
  \pr(\LENS|\saidLENS,T_k) &= \frac{1.0 \cdot p_0}
           {\left[ 1.0\cdot p_0 + 0.0\cdot(1 - p_0) \right]}
   &= 1.0,
\end{align}
as we might expect for such a {\it perfect} classifier.  Meanwhile, a
hypothetical volunteer who (for some reason) wilfully always submits the wrong
classification would have an agent with the column-swapped confusion matrix
\be
  \mathcal{M}^{\rm obtuse} = 
  \begin{bmatrix}
    1.0 & 0.0 \\
    0.0 & 1.0
  \end{bmatrix},
\ee
and would submit $C = \saidNOT$ for this subject. However, such a volunteer
would nevertheless be submitting useful information, since given the above
confusion matrix, their agent would calculate
\begin{align}
  \pr(\LENS|\saidNOT,T_k) &= \frac{1.0 \cdot p_0}
           {\left[ 1.0\cdot p_0 + 0.0\cdot(1 - p_0) \right]}
   &= 1.0.
\end{align}
{\it Obtuse} classifiers turn out to be as helpful as {\it perfect} ones.

\subsection{Information Contribution}
\label{appendix:swap:examples}

The information likely to be contributed by each agent for a given subject
can be estimated before the next classification of that subject is made, 
just from its confusion matrix.
The Shannon entropy generated by a classifier upon performing a
classification is
\be
  \langle S_k \rangle =  
   -P_{\rm right}\cdot\log_2{P_{\rm right}} -  P_{\rm wrong}\cdot\log_2P_{\rm wrong}\,,
\ee
where $P_{\rm right}$ and $P_{\rm wrong}$ are the averages of the diagonal and
the off-diagonal elements of the confusion matrix, respectively, and $\langle
S_k \rangle$ is measured in ``bits.'' These averages represent the probability
of a classifier to get a classification right or wrong, respectively. We define
the information contributed by a classifier as
\begin{equation}
I_k = 1 - S_k\,.
  \label{eq:app:info}
\end{equation}
\Eref{eq:app:info} gives the required result, that both the hypothetical {\it
perfect} and {\it obtuse} classifiers contribute 1 bit of information each, per
classification. Classifiers whose agent's confusion matrix is such that $P_{\rm
right}=P_{\rm wrong}=0.5$, contribute zero bits of information. Such users
identify a lens correctly with the same probability as they misclassify a dud
image to contain a lens, and thus their classification is of no value.

@@

\subsection{Updating the Subject Probabilities}
\label{appendix:swap:examples}

@@

\subsection{Uncertainty in the Agent Confusion Matrices}
\label{appendix:swap:uncertainty}

The confusion matrix obtained from the application of Equation
(\ref{eq:app:fraction}) has some inherent noise which reduces as the number of
training subjects classified by the agent's volunteer increases. For
simplicity, the discussion thus far assumed the case when the confusion matrix
is known perfectly; in practice, we allow for uncertainty in the agent
confusion matrices by averaging over a small number of samples drawn from
Binomial distributions characterised by the matrix elements
$\pr(C_k|\LENS,T_k)$ and  $\pr(C_k|\NOT,T_k)$. The associated standard
deviation in the estimated subject probability provides an error bar for this
quantity.

% For ease of notation, we will denote $\pr(C_k|\LENS,T_k)\equiv p_L$ and
% $\pr(C_k|\NOT,T_k)\equiv p_N$. In reality, there is a probability distribution
% for both $p_L$ and $p_N$. Let $p_0$ be the prior probability of the subject
% being a lens. Then the posterior probability, $p_0'$ of the subject being a
% lens after the classification $C_k$ is
% \be
%   \label{eq:app:sec}
% p_0' = \frac{p_L p_0}{\left[ p_L p_0 + p_N (1- p_0) \right]},
% \ee
% The posterior probability distribution $p_0'$ can be obtained by marginalizing
% over the probability distributions of $p_L$, $p_N$ and the prior probability
% distribution $p_0$ such that,
% \be
% P(p_0') = \int p_0' P(p_L) P(p_N) P(p_0) dp_L dp_N dp_0\,.
% \ee
% This marginalization is not analytically tractable. Therefore, we have
% implemented the following Monte-Carlo solution for this problem.
%
